---
title: "pksensi_NB"
output:
  html_document:
    df_print: paged
  html_notebook: default
---


```{r message = FALSE, warning=FALSE}
# Install the pksensi
# devtools::install_github("nanhung/pksensi")
library(pksensi) 
library(sensitivity)
```

# Use the add-in function (sobol) and do the testing
## The eFAST method in sensitivity package 
```{r message = FALSE}
x<-fast99(factors = 8, n = 400,
           q = "qunif", q.arg = list(min = 0, max = 1))
y<-sobol.fun(x$X)
tell(x, y)
x
plot(x)
```

## The eFAST with random phase shift method in pksensi package
```{r message = FALSE}
set.seed(1234)
x<-rfast99(factors = 8, n = 400,
           q = "qunif", q.arg = list(min = 0, max = 1), replicate = 20, conf = 0.95)
y<-solve_fun(x, model = pksensi:::sobol.fun)
tell2(x, y)
x
plot(x)
```

# Exercise: Flip-Flop Kinetics (Analytical approach)

```{r message = FALSE}
# Define Flip-Flop kinetic function
FFPK <- function(parameters, times, dose = 320){
  A <- (dose * parameters[1])/( parameters[3]*( parameters[1]- parameters[2]))
  CONC <- A*exp(- parameters[2] * times) - A*exp(- parameters[1] * times)
  return(CONC)
}

# Define parameter distribution
q = "qunif"
q.arg = list(list(min = 0.5, max = 1.5), 
             list(min = 0.02, max = 0.3),
             list(min = 20, max = 60))

set.seed(1234)
x<-rfast99(factors=c("KA","KE","V"),
           n = 400, q = q, q.arg = q.arg, rep = 20, conf = 0.95)
```


```{r message = FALSE}
# Solve the function
times <- seq(from = 0.25, to = 24.25, by = 0.5)
y<-solve_fun(x, model = FFPK, times = times)
tell2(x,y)
x
plot(x)
```

# HTTK (Numerical approach)
# Case 1: 1 Compartment model

# Model compiled
```{r message = FALSE}
##### MCSim under R (use deSolve package)
# pros: Don't need to create in file
# cons: Slower than in pure MCSim

compile <- function (mName, model = F) {
  if (model == T){ # Generate the ".c" file and "_inits.R" from model file 
    if(file.exists(paste0(mName, ".model")) && .Platform$OS.type == "unix"){
      system (paste0("mod -R ", mName, ".model ", mName,".c")) # model to c file
    } else if (.Platform$OS.type == "windows") {
      warning("The model compiled function haven't supprot Windows system")
    }
  }
  if (is.loaded("derivs", PACKAGE=mName))
    dyn.unload(paste0(mName,.Platform$dynlib.ext))
  
  if (.Platform$OS.type == "windows") {
    if (!(devtools::find_rtools() == TRUE)) {
      # The windows user neet to install Rtools to complile the c file
      # The MacOS user neet to install Xcode
      warning("The Rtools should be installed first") 
    }
  } 
  system (paste0("R CMD SHLIB ", mName, ".c")) # create .o and .so (or .dll) files
  
  dyn.load(paste0(mName, .Platform$dynlib.ext))
  
  if(file.exists(paste0(mName, "_inits.R"))){
    source(paste0(mName, "_inits.R"))
  }
}
```

# Compile the c file
```{r message = FALSE}
mName = "pbtk1comp"
compile(mName) 
# The compile function can generate the R file for input parameter and initial state 
```



```{r message = FALSE}
times <- seq(from = 0.5, to = 24.5, by = 1)
parameters <- initparms1comp()
initState <- initState1comp(parms=parameters)
initState[1] <- 10

parameters
initState

```


```{r message = FALSE}
library(httk)
params <- parameterize_1comp(chem.name = "theophylline")
params
```

```{r message = FALSE}
# Use httk generated parameter values with 50% uncertainty
LL <- 0.5
UL <- 1.5

q = "qunif"
q.arg = list(list(min = params$Vdist * LL, max = params$Vdist * UL),
             list(min = params$kelim * LL, max = params$kelim * UL),
             list(min = params$kgutabs * LL, max = params$kgutabs * UL))

set.seed(1234)
x<-rfast99(factors = c("vdist", "ke", "kgutabs"),
           n = 100, q = q, q.arg = q.arg, rep = 10, conf = 0.99)

# Use deSolve to solve ode (take some time)
y<-solve_fun(x, times, parameters = parameters, initState, outnames = "Ccompartment",
             dllname = mName, func = "derivs1comp", initfunc = "initmod1comp")
tell2(x,y)
```


```{r message = FALSE}
pksim(y, log = T) # Use log-transformed scale
points(Theoph$Time, log(Theoph$conc), col=Theoph$Subject, pch=19)
```

```{r message = FALSE}
check(x)
plot(x, cut.off = 0.05)
```

# Case 2: 3 CompartmentPBPK model
```{r message = FALSE}
mName = "3compPBPKmodel"
compile(mName)

parameters <- initparms3comp()
initState <- initState3comp(parms=parameters)
initState[1] <- 10

parameters
initState

```

```{r message = FALSE}
params <- httk::parameterize_3comp(chem.name = "theophylline")
params
```

```{r message = FALSE}
# 20% uncertainty
LL <- 0.8
UL <- 1.2

# 20 parameters
q = "qunif"
q.arg = list(list(min = params$BW * LL, max = params$BW * UL),
             list(min = params$Clmetabolismc * LL, max = params$Clmetabolismc * UL),
             list(min = params$kgutabs * LL, max = params$kgutabs * UL),
             list(min = params$Qcardiacc * LL, max = params$Qcardiacc * UL),
             list(min = params$Qgfrc * LL, max = params$Qgfrc * UL),
             list(min = params$Qgutf * LL, max = params$Qgutf * UL),
             list(min = params$Qliverf * LL, max = params$Qliverf * UL),
             list(min = params$Vgut * LL, max = params$Vgut * UL),
             list(min = params$Vliver * LL, max = params$Vliver * UL),
             list(min = params$Vrest * LL, max = params$Vrest * UL),
             list(min = params$Funbound.plasma * LL, max = params$Funbound.plasma * UL),
             list(min = params$Clmetabolism * LL, max = params$Clmetabolism * UL),
             list(min = params$Qcardiac * LL, max = params$Qcardiac * UL),
             list(min = params$Qgfr * LL, max = params$Qgfr * UL),
             list(min = params$Qgut * LL, max = params$Qgut * UL),
             list(min = params$Qliver * LL, max = params$Qliver * UL),
             list(min = params$Kliver2pu * LL, max = params$Kliver2pu * UL),
             list(min = params$Krest2pu * LL, max = params$Krest2pu * UL),
             list(min = params$Kgut2pu * LL, max = params$Kgut2pu * UL),
             list(min = params$Rblood2plasma * LL, max = params$Rblood2plasma * UL))

factors <- names(parameters)

#x<-rfast99(factors = factors, n = 2000, q = q, q.arg = q.arg, rep = 10, conf = 0.95)
#x<-rfast99(factors = factors, n = 4000, q = q, q.arg = q.arg, rep = 10, conf = 0.95)

times <- seq(from = 0.5, to = 24.5, by = 1)

# Takes about 1 - 2 hour for the computation
# y<-solve_fun(x, times, parameters = parameters, initState, outnames = "Crest",
#              dllname = mName, func = "derivs3comp", jacfunc = "jac3comp", initfunc = "initmod3comp")
# tell2(x,y)
```

# Sample number n = 2000
```{r message = FALSE}
# sample number = 2000
load(file = "3comp_2000.rda") # parameter space
load(file = "3comp_2000y.rda") # model output
cat(file.size(file = "3comp_2000.rda")/1e6, "MB", "\n") 
cat(file.size(file = "3comp_2000y.rda")/1e6, "MB") 
```

```{r message = FALSE}
pksim(y, log = T)
points(Theoph$Time, log(Theoph$conc), col=Theoph$Subject, pch=19)
```

```{r message = FALSE}
plot(x, cut.off = 0.05)
check(x)
```

# Heapmap check
```{r message = FALSE}
heat_check(x, index = "SI") 
heat_check(x, index = "SI", order = T)
heat_check(x, index = "CI")
```

# Sample number n = 4000
```{r message = FALSE}
# sample number = 4000
load(file = "3comp_4000.rda")
load(file = "3comp_4000y.rda")
cat(file.size(file = "3comp_4000.rda")/1e6, "MB", "\n")
cat(file.size(file = "3comp_4000y.rda")/1e6, "MB")
```

```{r message = FALSE}
plot(x, cut.off = 0.05)
heat_check(x, index = "SI") 
heat_check(x, index = "SI", order = T)
heat_check(x, index = "CI")
```

#Customize heatmap

```{r message = FALSE}
# specific index (the default show all results)
heat_check(x, filter = c("first order","interaction"))
```

```{r message = FALSE}
# Continous
heat_check(x, filter = c("first order","total order"), category = F)
# Add value
heat_check(x, filter = c("first order","total order"), category = F, text = T)
heat_check(x, filter = c("first order","total order"), order = T, text = T)
```

# Customize color 
```{r message = FALSE}
# grey scale
heat_check(x) + ggplot2::scale_fill_grey(start = .9, end = .0)
heat_check(x, category = F) + ggplot2::scale_fill_gradient(low = "white", high = "black")
# viridis
heat_check(x) + viridis::scale_fill_viridis(discrete=TRUE)
heat_check(x, category = F) + viridis::scale_fill_viridis()

#
heat_check(x, filter = c("first order","interaction"), text = T, category = F) + ggplot2::scale_fill_gradient(low = "white", high = "blue") + ggplot2::ggtitle("The eFAST with random phase shift")
```