---
title: "pksensi_NB"
output:
  html_document:
    df_print: paged
  html_notebook: false
---

This [R Markdown](http://rmarkdown.rstudio.com) Notebook shows the main functions in [pksensi](https://github.com/nanhung/pksensi) package.

**pksensi** implements the global sensitivity analysis workflow to investigate the parameter uncertainty and sensitivity in pharmacokinetic (PK) models, especially the physiologically based pharmacokinetic (PBPK) model and advanced compartment absorption and transit (ACAT) model with multivariate output. The package also provide some functions to check the convergence and sensitivity of model parameters.

```{r message = FALSE, warning=FALSE}
# Install the pksensi, if the package haven't already:
# devtools::install_github("nanhung/pksensi")

# Import the pksensi
library(pksensi) 
```

## sensitivity::fast99 and pksensi::rfast99
Use the add-in function (sobol) and do the testing. The eFAST method in sensitivity package is used to compare the result from pksensi.

```{r message = FALSE}

library(sensitivity)
x<-fast99(factors = 8, n = 400,
           q = "qunif", q.arg = list(min = 0, max = 1))
y<-sobol.fun(x$X)
tell(x, y)
x
plot(x)
```

Here shows the eFAST with random phase shift method in pksensi package

```{r message = FALSE}
set.seed(1234)
x<-rfast99(factors = 8, n = 400,
           q = "qunif", q.arg = list(min = 0, max = 1), replicate = 20, conf = 0.95)
y<-solve_fun(x, model = pksensi:::sobol.fun) 
tell2(x, y) # like sensitivity::tell function, the tell2 function is used to combine the output result and model output.
x
plot(x)
```

**Note**
* The sole_fun can use to solve the function in analytical and numerical approach.

## Exercise: Flip-Flop Kinetics (Analytical approach)

```{r message = FALSE}
# Define Flip-Flop kinetic function
FFPK <- function(parameters, times, dose = 320){
  A <- (dose * parameters[1])/( parameters[3]*( parameters[1]- parameters[2]))
  CONC <- A*exp(- parameters[2] * times) - A*exp(- parameters[1] * times)
  return(CONC)
}

# Define parameter distribution
q = "qunif"
q.arg = list(list(min = 0.5, max = 1.5), 
             list(min = 0.02, max = 0.3),
             list(min = 20, max = 60))

# The replication and confidence interval are set to 20, 0.95, respectively.
set.seed(1234)
x<-rfast99(factors=c("KA","KE","V"),
           n = 400, q = q, q.arg = q.arg, rep = 20, conf = 0.95)
```

Unlike the sobol function, we need to define the time step here.

```{r message = FALSE}
times <- seq(from = 0.25, to = 24.25, by = 0.5)
y<-solve_fun(x, model = FFPK, times = times)
tell2(x,y)
x # print the time-dependent output of sensitivity index and convergence index to the console 
plot(x) # Visualize the printed result 
```

We can use log-tranformed parameter to adjust the large-scale in parameter sampling.

```{r message = FALSE}
# Define parameter distribution
q = "qunif"
q.arg = list(list(min = log(0.5), max = log(1.5)), 
             list(min = log(0.02), max = log(0.3)),
             list(min = log(20), max = log(60)))

set.seed(1234)
x<-rfast99(factors=c("KA","KE","V"),
           n = 400, q = q, q.arg = q.arg, rep = 20, conf = 0.95)
```

Use `lnparam = T` to let `solve_fun` transfer the input parameter and estimate the ourput. Therefore, we don't need to add the log-transormed parameter in model file.

```{r message = FALSE}
# Solve the function
times <- seq(from = 0.25, to = 24.25, by = 0.5)
y<-solve_fun(x, model = FFPK, times = times, lnparam = T)
tell2(x,y)
x
plot(x)
```

The `pksim` function is use to simulate the output result. This method can use be used to compare the experimental data.  

```{r message = FALSE}
pksim(y)
points(Theoph$Time, Theoph$conc, col=Theoph$Subject, pch=19)
```

The `check` function can summarize the sensitivity index and convergence index. The `SI` and `CI` can change to other value.

```{r message = FALSE}
check(x)
check(x, SI = 0.05, CI = 0.05)
plot(x, cut.off = 0.05)
```

## HTTK (Numerical approach)
### Case 1: One-compartment model

* Model compiled (under development)
We test One-compartment and 3-compartment PBPK model in this section. The c file was sourced from httk pckage.

The c file need to compile to executable file (dynamic-link libraries (.dll) on Windows or shared objects (.so) on unix systems).

The windows user neet to install [Rtools](https://cran.r-project.org/bin/windows/Rtools/) to complile the c file. The MacOS user neet to install Xcode (https://developer.apple.com/). The GNU compilers are included by default in most Linux distributions.

The follwong function had been test in Windows 7, Ubuntu 18.04, and MacOS 10.13.4.

```{r message = FALSE}
##### MCSim under R (use deSolve package)
# pros: Don't need to create in file
# cons: Slower than in pure MCSim

compile <- function (mName, model = F) {
  if (model == T){ # Generate the ".c" file and "_inits.R" from model file 
    if(file.exists(paste0(mName, ".model")) && .Platform$OS.type == "unix"){
      system (paste0("mod -R ", mName, ".model ", mName,".c")) # model to c file
    } else if (.Platform$OS.type == "windows") {
      warning("The model compiled function haven't supprot Windows system")
    }
  }
  if (is.loaded("derivs", PACKAGE=mName))
    dyn.unload(paste0(mName,.Platform$dynlib.ext))
  
  if (.Platform$OS.type == "windows") {
    if (!(devtools::find_rtools() == TRUE)) {
      warning("The Rtools should be installed first") 
    }
  } 
  system (paste0("R CMD SHLIB ", mName, ".c")) # create .o and .so (or .dll) files
  
  dyn.load(paste0(mName, .Platform$dynlib.ext))
  
  if(file.exists(paste0(mName, "_inits.R"))){
    source(paste0(mName, "_inits.R"))
  }
}
```

Compile the c file

```{r message = FALSE}
mName = "pbtk1comp"
compile(mName) 
# The compile function can generate the R file for input parameter and initial state 
```

Check the file

```{r message = FALSE}
file.exists("pbtk1comp.so")
```

Use `pbtk1comp_inits.R` file to define the parameter and initial state.

```{r message = FALSE}
times <- seq(from = 0.5, to = 24.5, by = 1)
parameters <- initparms1comp()
initState <- initState1comp(parms=parameters)
initState[1] <- 10
outnames <- Outputs1comp
parameters
initState

```

Import [httk](https://cran.r-project.org/web/packages/httk/index.html)to get the parameter value for specific chamical.

```{r message = FALSE}
library(httk)
params <- parameterize_1comp(chem.name = "theophylline")
params
```

Set the parameter uncertainty and solve ODE by R [deSolve](https://cran.r-project.org/web/packages/deSolve/index.html) package.

```{r message = FALSE}
# Use httk generated parameter values with 50% uncertainty
LL <- 0.5
UL <- 1.5

q = "qunif"
q.arg = list(list(min = params$Vdist * LL, max = params$Vdist * UL),
             list(min = params$kelim * LL, max = params$kelim * UL),
             list(min = params$kgutabs * LL, max = params$kgutabs * UL))

set.seed(1234)
x<-rfast99(factors = c("vdist", "ke", "kgutabs"),
           n = 400, q = q, q.arg = q.arg, rep = 10, conf = 0.99)

# Use deSolve to solve ode (take some time)
y<-solve_fun(x, times, parameters = parameters, initState, outnames = outnames,
             dllname = mName, func = "derivs1comp", initfunc = "initmod1comp", output = "Ccompartment")
tell2(x,y)
```

Here use n = 400 to reach converge.


```{r message = FALSE}
plot(x)
pksim(y, log = T) # Use log-transformed scale
points(Theoph$Time, log(Theoph$conc), col=Theoph$Subject, pch=19)
```


## Case 2: The 3-CompartmentPBPK model


```{r message = FALSE}
mName = "3compPBPKmodel"
compile(mName)

parameters <- initparms3comp()
initState <- initState3comp(parms=parameters)
initState[1] <- 10
outnames <- Outputs3comp

parameters
initState

```

Define theparameter value

```{r message = FALSE}
params <- httk::parameterize_3comp(chem.name = "theophylline")
params
```

Set the parameter uncertainty and solve ODE

```{r message = FALSE}
# 20% uncertainty
LL <- 0.8
UL <- 1.2

# 20 parameters
q = "qunif"
q.arg = list(list(min = params$BW * LL, max = params$BW * UL),
             list(min = params$Clmetabolismc * LL, max = params$Clmetabolismc * UL),
             list(min = params$kgutabs * LL, max = params$kgutabs * UL),
             list(min = params$Qcardiacc * LL, max = params$Qcardiacc * UL),
             list(min = params$Qgfrc * LL, max = params$Qgfrc * UL),
             list(min = params$Qgutf * LL, max = params$Qgutf * UL),
             list(min = params$Qliverf * LL, max = params$Qliverf * UL),
             list(min = params$Vgut * LL, max = params$Vgut * UL),
             list(min = params$Vliver * LL, max = params$Vliver * UL),
             list(min = params$Vrest * LL, max = params$Vrest * UL),
             list(min = params$Funbound.plasma * LL, max = params$Funbound.plasma * UL),
             list(min = params$Clmetabolism * LL, max = params$Clmetabolism * UL),
             list(min = params$Qcardiac * LL, max = params$Qcardiac * UL),
             list(min = params$Qgfr * LL, max = params$Qgfr * UL),
             list(min = params$Qgut * LL, max = params$Qgut * UL),
             list(min = params$Qliver * LL, max = params$Qliver * UL),
             list(min = params$Kliver2pu * LL, max = params$Kliver2pu * UL),
             list(min = params$Krest2pu * LL, max = params$Krest2pu * UL),
             list(min = params$Kgut2pu * LL, max = params$Kgut2pu * UL),
             list(min = params$Rblood2plasma * LL, max = params$Rblood2plasma * UL))

factors <- names(parameters)
times <- seq(from = 0.5, to = 24.5, by = 1)

#x<-rfast99(factors = factors, n = 4000, q = q, q.arg = q.arg, rep = 10, conf = 0.95)
#y<-solve_fun(x, times, parameters = parameters, initState, outnames = outnames,
#             dllname = mName, func = "derivs3comp", initfunc = "initmod3comp", 
#             output = "Crest")
# tell2(x,y)
```

# Sample number n = 2000
```{r message = FALSE}
# sample number = 2000
load(file = "3comp_2000.rda") # parameter space
load(file = "3comp_2000y.rda") # model output
cat(file.size(file = "3comp_2000.rda")/1e6, "MB", "\n") 
cat(file.size(file = "3comp_2000y.rda")/1e6, "MB") 
```

```{r message = FALSE}
pksim(y, log = T)
points(Theoph$Time, log(Theoph$conc), col=Theoph$Subject, pch=19)
```

```{r message = FALSE}
plot(x, cut.off = 0.05)
check(x)
```

# Heapmap check
```{r message = FALSE}
heat_check(x, index = "SI") 
heat_check(x, index = "SI", order = T)
heat_check(x, index = "CI")
```

# Sample number n = 4000
```{r message = FALSE}
# sample number = 4000
load(file = "3comp_4000.rda")
load(file = "3comp_4000y.rda")
cat(file.size(file = "3comp_4000.rda")/1e6, "MB", "\n")
cat(file.size(file = "3comp_4000y.rda")/1e6, "MB")
```

```{r message = FALSE}
plot(x, cut.off = 0.05)
heat_check(x, index = "SI") 
heat_check(x, index = "SI", order = T)
heat_check(x, index = "CI")
```

## Case 3: ACAT-PBPK model model

The ACAT_like.model was sourced from 'mcsim-6.0.0/examples/ACAT-PBPK'
```{r message = FALSE}
file.exists("ACAT_like.model")
```

First, compile the MCSim model file to the executable file
The model file will generate the C model file 'ACAT_like.c' to linke with R deSolve package
It can also create the R parameter initialization file 'ACAT_like_inits.R'
The executable file 'ACAT_like.so' or 'ACAT_like.dll' will generate in the compile process.
```{r message = FALSE}
source("compile.R")
mName <- "ACAT_like"
compile(mName, model = T)
```

Seting Initial state and parameter value. Al parameter value can be check by using `initParms()`. Also, the `initStates()` can use to check all state conditions. The `Outputs` included all variables that were defined in the statement of outputs in 'ACAT_like.model'
```{r message = FALSE}
newState <- c(A_stom_lu = 1780) # Ingestion dose uM
newParms <- c(BDM = 70, # Body weight (kg)
              MM = 180.17, # Molecular weight
              f_Abs_jeju  = 1, # only cosider the absorption in jejunum 
              Peff = 0.175, # Effective permeability of g.i. tract epithelia
              Ratio_BP = 0.824, # Plasma / blood concentration ratio
              # Tissue / blood partition coefficients
              PC_adip   = 1, # adipose tissue
              PC_kid    = 1, # Kidney
              PC_rpt    = 1, # richly perfused tissues (viscera)
              PC_ppt    = 1, # poorly perfused tissues (muscle and skin)
              PC_liver  = 1, # liver
              PC_stom   = 1, # stomach 
              PC_duod   = 1, # duodenum
              PC_jeju   = 1, # jejunum
              PC_ileon  = 1, # ileon
              PC_cecum  = 1, # cecum
              PC_colon  = 1, # colon
              # Fraction unbound in compartments
              Fu_adip     = 1,
              Fu_kid      = 1,
              Fu_ppt      = 1,
              Fu_rpt      = 1,
              Fu_blood    = 1,
              Fu_liver    = 1,
              Fu_portvein = 1,
              Fu_stom     = 1,
              Fu_duod     = 1,
              Fu_jeju     = 1,
              Fu_ileon    = 1,
              Fu_cecum    = 1,
              Fu_colon    = 1,
              Fu_plasma   = 0.016,
              Vmax_vitro = 2.48, # Michaelis-Menten Vmax (micromol/min/mg microsomal proteins)
              Km_vitro = 3.31, # Michaelis-Menten Km (microM)
              Kle_kid = 0.085) # Renal elimination clearance (L/h)
parameters <- initParms(newParms=newParms)
initState <- initStates(newStates=newState)
outnames <- Outputs
length(parameters); round(parameters, digits = 4)
length(initState); initState
length(outnames); outnames
```

Define the parameter distribution (30 parameters)
```{r message = FALSE}
factors <- c("Peff",
             "Ratio_BP",
             "PC_adip",
             "PC_kid",
             "PC_rpt",
             "PC_ppt",
             "PC_liver",
             "PC_stom",
             "PC_duod",
             "PC_jeju",
             "PC_ileon",
             "PC_cecum",
             "PC_colon",
             "Fu_adip",
             "Fu_kid",
             "Fu_ppt",
             "Fu_rpt",
             "Fu_blood",
             "Fu_liver",
             "Fu_portvein",
             "Fu_stom",
             "Fu_duod",
             "Fu_jeju",
             "Fu_ileon",
             "Fu_cecum",
             "Fu_colon",
             "Fu_plasma",
             "Vmax_vitro", 
             "Km_vitro",
             "Kle_kid")
q = "qunif"
# Use log-transformed parameter value
q.arg = list(list(min = -2.3, max = 0), # Peff
             list(min = -2.3, max = 0), # Ratio_BP
             list(min = -4.6, max = 4.6), #PC_adip
             list(min = -4.6, max = 4.6), #PC_kid
             list(min = -4.6, max = 4.6), #PC_rpt
             list(min = -4.6, max = 4.6), #PC_ppt
             list(min = -4.6, max = 4.6), #PC_liver
             list(min = -4.6, max = 4.6), #PC_stom
             list(min = -4.6, max = 4.6), #PC_duod
             list(min = -4.6, max = 4.6), #PC_jeju
             list(min = -4.6, max = 4.6), #PC_ileon
             list(min = -4.6, max = 4.6), #PC_cecum
             list(min = -4.6, max = 4.6), #PC_colon
             list(min = -4.6, max = 0), # Fu_adip
             list(min = -4.6, max = 0), # Fu_kid
             list(min = -4.6, max = 0), # Fu_ppt
             list(min = -4.6, max = 0), # Fu_rpt
             list(min = -4.6, max = 0), # Fu_blood
             list(min = -4.6, max = 0), # Fu_liver
             list(min = -4.6, max = 0), # Fu_portvein
             list(min = -4.6, max = 0), # Fu_stom
             list(min = -4.6, max = 0), # Fu_duod
             list(min = -4.6, max = 0), # Fu_jeju
             list(min = -4.6, max = 0), # Fu_ileon
             list(min = -4.6, max = 0), # Fu_cecum
             list(min = -4.6, max = 0), # Fu_colon
             list(min = -4.6, max = 0), # Fu_plasma
             list(min = 0, max = 4.6), # Vmax_vitro
             list(min = 0, max = 4.6), # Km_vitro
             list(min = -4.6, max = 0)) # Kle_kid
```

Set the time step and input the parameters in `rfast99`
```{r eval=FALSE}
times <- seq(from = 0.5, to = 12.5, by = 1)
x<-rfast99(factors = factors, n = 2000, q = q, q.arg = q.arg, rep = 20, conf = 0.95) 
#y<-solve_fun(x, times, parameters = parameters, initState = initState, 
#             outnames = outnames, dllname = mName,
#             func = "derivs", initfunc = "initmod", lnparam = T, output = "C_blood")
# user   system  elapsed 
# 3801.957    6.150 3808.827 
cat(file.size(file = "acat_2000.rda")/1e6, "MB", "\n")
cat(file.size(file = "acat_2000y.rda")/1e6, "MB")
load(file = "acat_2000.rda")
load(file = "acat_2000y.rda")
```

Create the basic plot
```{r fig.height=9, fig.width=11, message=FALSE}
pksim(y)
plot(x,cut.off = 0.05)
```

Check sensitivity parameter
```{r fig.height=6, fig.width=8, message=FALSE}
heat_check(x, order =T)
heat_check(x, order =T, category = F, text= T)
heat_check(x, index = "CI")
```

## Customize heatmap

```{r message = FALSE}
# specific index (the default show all results)
heat_check(x, filter = c("first order","interaction"))
```

```{r message = FALSE}
# Continous
heat_check(x, filter = c("first order","total order"), category = F)
# Add value
heat_check(x, filter = c("first order","total order"), category = F, text = T)
heat_check(x, filter = c("first order","total order"), order = T, text = T)
```

## Customize color 
```{r message = FALSE}
# grey scale
heat_check(x) + ggplot2::scale_fill_grey(start = .9, end = .0) + ggplot2::ggtitle("Grey scale (discrete)")
heat_check(x, category = F) + ggplot2::scale_fill_gradient(low = "white", high = "black") + ggplot2::ggtitle("Grey scale (continue)")

# viridis for color blindness
heat_check(x) + viridis::scale_fill_viridis(discrete=TRUE) + ggplot2::ggtitle("viridis for color blindness")
heat_check(x, category = F) + viridis::scale_fill_viridis() + ggplot2::ggtitle("viridis for color blindness")

#
heat_check(x, filter = c("first order","interaction"), text = T, category = F) + ggplot2::scale_fill_gradient(low = "white", high = "blue") + ggplot2::ggtitle("The eFAST with random phase shift")
```